---
title: "Post 3 - How to use parallel iteration to create plots"
description: |
  Examples of using pmap() to loop through your data and generate a series of plots.  
author:
  - name: Joanna Wright
    url: https://s-gluck.github.io/funprog_blog
date: 06-01-2020
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 2
---

This post describes examples of using iteration to loop through a data set and plot.

# Set-up

```{r setup }
knitr::opts_chunk$set(echo = TRUE)

```

<aside>
Libraries for this post
</aside>
```{r}
library(ggplot2)
library(tidyverse)
library(janitor)
library(glue)
library(purrr)
```


## Loading the data 
For details on how to load the data, see [Post #1 by Brock](https://s-gluck.github.io/funprog_blog/posts/2020-05-11-blog_post_one/).
For details about this dataset, see [Post #2 by Akhila and Stephanie](https://s-gluck.github.io/funprog_blog/posts/2020-05-08-stephaniepost/). 

```{r loading the data, results='hide'}

files <- list.files(here::here("data"),
                    full.names = TRUE)
files
d <- read_csv(files[1]) %>%
  clean_names()

```

# Nesting and using `pmap()` for plotting 

## Nesting
Let's say we want to create a separate plot for mobility trends in each country.  First, we need to `group_by` country and `nest()` the rest of the data. This creates a list of tibbles, one for each country. I've called `reactable()` here so that we can return a simple, interactive table to see the results of nesting. 

```{r}

d %>%
  group_by(country) %>% 
  nest() %>%
  reactable::reactable()

```
Now we can use this as the input for our plots.

## pmap() and ggplot() for generating a series of plots

### About `pmap()`: pmap(.l, .f, ...)
`pmap(.l, .f, ...)` is a variation of `map()` which allows us to apply a function to multiple vectors simultaneously. According to the [help documentation](https://purrr.tidyverse.org/reference/map2.html), it takes 2 or more arguments: 

* .d = a list of vectors (such as a data frame).
* .f = 	A function, formula, or vector (not necessarily atomic). *In the example below, we've supplied another vector.*
* ... = Additional arguments passed on to the mapped function. **This is what makes `pmap()` so flexible!** 

## Getting started with plotting 
In the syntax  below, you'll see that the first vector listed corresponds to ..1 (in this case, country) and the second vector listed (data) corresponds to ..2. The vectors listed in `aes()` are from within the data, which we specifed with ..2. For this example, we'll create one plot for each country, for just one of the mobility categories (workplaces). 

```{r parallel iteration, workplaces mobility trends for each country}

plots_workplaces <- d %>%
  group_by(country) %>% 
  nest() %>% 
  mutate(plots = pmap(list(country, data),  
                      ~ggplot(..2, aes(date, workplaces)) +
                        geom_point() + 
                        geom_line()))

```

Test to see what the plots look like: 

```{r}

plots_workplaces$plots[1]
plots_workplaces$plots[2]

# or we could view all of them like this: 
# plots_workplaces$plots[1:19]

```

## Plot improvements and customization 
We can add a unique title to each plot with `{..1}` to select the 1st column (country) to insert country names into the title.

Plus a few other visualization adjustments... 

```{r adding unique title to each plot with country name}

plots_workplaces <- d %>%
  group_by(country) %>% 
    nest() %>%
    mutate(plots = pmap(list(country, data),
    ~ggplot(..2, aes(date, workplaces)) +
      geom_line(color = "grey70") +
      geom_point(aes(color = workplaces > 0), 
                 size = 2) +
      scale_color_brewer(palette = "Set2") +      
      theme(axis.title.x = element_blank(), 
            legend.position = 'none', 
            plot.background = element_rect(fill="gray20"),
            panel.background = element_rect(fill="gray20"),
            panel.grid.minor = element_blank(), 
            plot.title = element_text(color="white",
                                      hjust= 0,
                                      vjust=1, 
                                      size=rel(1.5)), 
            axis.text = element_text(color="white", 
                               size=rel(1.5)),
            axis.text.y  = element_text(hjust=1), 
            axis.title.y = element_text(color = "white"),
            plot.caption = element_text(color = "white")) +
      labs(title = glue("Mobility trends to/from workplaces in {..1}"), #unique title
           y = "% change in mobility", 
           caption = "Googlemaps data from https://www.kaggle.com/gustavomodelli/covid-community-measures"))
    )

# testing: 
plots_workplaces$plots[1]
plots_workplaces$plots[2]

# or we could see all of them: 
# plots_workplaces$plots[1:19]
```

# Another application of pmap() 
What if we wanted to see plots for *all* mobility categories, not just workplaces? We can apply the same logic as above, but first nest by both country *and* mobility category. 

In order to do so, we'll first have to restructure the data so that mobility category is it's own column. This can be done using `pivot_longer()`: 

```{r}

by_country_mobil_cat <- d %>%
  pivot_longer(
    cols = 4:9,
    names_to = "mobil_category",
    values_to = "perc_change"
  ) %>%
  nest(-country, -mobil_category)

#head(by_country_mobil_cat)

reactable::reactable(by_country_mobil_cat)

```

Now we can use the same code structure as before for plotting. `pmap()` is super useful here because we can `list()` any number of columns. Note that here the first argument for `ggplot()` is now `..3`, referring to the 3rd item ("data") in the `pmap(list())`. Also, the `aes()` now calls for perc_change instead of parks or residential, etc., because that is the name of the column after using pivot_longer. 

```{r}

all_plots <- by_country_mobil_cat %>% 
  mutate(plots = pmap(list(country, mobil_category, data),
                      ~ggplot(..3, aes(date, perc_change)) +
      geom_line(color = "grey70") +
      geom_point(aes(color = perc_change > 0), 
                 size = 2) +
      scale_color_brewer(palette = "Set2") +      
      theme(axis.title.x = element_blank(), 
            legend.position = 'none', 
            plot.background = element_rect(fill="gray20"),
            panel.background = element_rect(fill="gray20"),
            panel.grid.minor = element_blank(), 
            plot.title = element_text(color="white",
                                      hjust= 0,
                                      vjust=1, 
                                      size=rel(1.5)), 
            axis.text = element_text(color="white", 
                               size=rel(1.5)),
            axis.text.y  = element_text(hjust=1), 
            axis.title.y = element_text(color = "white"),
            plot.caption = element_text(color = "white")) +
      labs(title = glue("Mobility trends to/from {..2} in {..1}"), #unique title
           y = "% change in mobility", 
           caption = "Googlemaps data from https://www.kaggle.com/gustavomodelli/covid-community-measures"))
    )

all_plots$plots[1]
all_plots$plots[10]
all_plots$plots[20]


```

# Using `walk2()` to save plots 

We can use walk2() to save our plots. It's similar to map() but doesn't print anything to the screen, instead creating files and file pathways to save to. 

```{r eval=FALSE}

fs::dir_create(here::here("plots", "mobility-plots")) # creates a folder called "plots" and a subdirectory called "mobility"

# then create a vector that has all of the file paths: 

files_c <- str_replace_all(tolower(all_plots$country), " ", "-")
files_m <- str_replace_all(all_plots$mobil_category, "_", "-")
paths <- here::here("plots", "mobility-plots", glue("{files_c}-{files_m}.png"))
#paths

walk2(paths, all_plots$plots, ggsave,
      width = 9.5, 
      height = 6.5,
      dpi = 500)

```



