---
title: "Post 3"
description: |
  Using iteration and writing functions to create plots. 
author:
  - name: Joanna Wright
    url: https://example.com/norajones
date: 05-08-2020
output:
  distill::distill_article:
    self_contained: false
---

This post describes examples of using iteration, parallel iteration, and writing functions in order to loop through a data set to extract information and create plots.

## Set-up

```{r setup }
knitr::opts_chunk$set(echo = TRUE)

```

<aside>
Libraries for this post
</aside>
```{r}
library(ggplot2)
library(tidyverse)
library(janitor)
library(glue)
library(purrr)
```


Loading the data. For details, see [Brock's post](link here).

```{r loading the data, results='hide'}

files <- list.files(here::here("data"),
                    full.names = TRUE)
files
d <- read_csv(files[1]) %>%
  clean_names()

```

## Task 1: Plotting mobility trends by country 

Let's say we want to create a separate plot for mobility trends in each country.  First, we need to `group_by` country and `nest()` the rest of the data. This creates a list of tibbles, one for each country. I've called `head()` here so that we can see what the first few rows look like. 

```{r}

d %>%
  group_by(country) %>% 
  nest() %>%
  head()

```

Now we can use this as the input for our plots. `pmap()` is a variation of `map()` which allows us to apply a function to multiple vectors simultaneously. In `pmap()`, the first vector listed corresponds to ..1 (in this case, country) and the second vector listed (data) corresponds to ..2. The vectors listed in `aes()` are from within the data, which we specifed with ..2. For this example, we'll create one plot for each country, for just one of the mobility categories - workplaces. 

```{r parallel iteration, workplaces mobility trends for each country}

plots_workplaces <- d %>%
  group_by(country) %>% 
  nest() %>% 
  mutate(plots = pmap(list(country, data),  
                      ~ggplot(..2, aes(date, workplaces)) +
                        geom_point() + 
                        geom_line()))

```

Test to see what the plots look like: 
```{r}

plots_workplaces$plots[1]
plots_workplaces$plots[2]

```

We could view all of them with: 

```{r eval=FALSE}

plots_workplaces$plots[1:19]

```

Let's tidy up this plot a bit more.  
To add a unique title to each plot, use `{..1}` to select the 1st column, country, to insert country names into the title. 

```{r adding unique title to each plot with country name}

# adding unique title: 
plots_workplaces1 <- d %>%
  group_by(country) %>% 
    nest() %>%
    mutate(plots = pmap(list(country, data),
    ~ggplot(..2, aes(date, workplaces)) +
      geom_line() +
      geom_point() + 
      theme(axis.title.x = element_blank()) +
      labs(title = glue("Mobility trends for workplaces in {..1}"), #unique title
           y = "% change"))
    )

# testing: 
plots_workplaces1$plots[1]
plots_workplaces1$plots[2]

# or we could see all of them: 
# plots_workplaces1$plots[1:19]
```

So far, we've looked at mobility trends in one particular category (workplaces), for each country. But what if we wanted to quickly see mobility trends for each of the columns, not just workplaces?

To do this, we need to nest as we did before, but this time by both country and mobility category. To do so, we first need to restructure the data so that we have a column for mobility category and a column for percent change: 

```{r}

by_country_mobil_cat <- d %>%
  pivot_longer(
    cols = 4:9,
    names_to = "mobil_category",
    values_to = "perc_change"
  ) %>%
  nest(-country, -mobil_category)

head(by_country_mobil_cat)

```

Then, we can generalize our plot with a function, which we could then apply to any selection of mobility categories.  

Our function will have 2 arguments: the data frame, and the mobility category. The x-axis variable, time, will remain the same, so it is supplied in the body of the function rather than as an argument. 

```{r}

# Now generalize to a function: 
create_plot <- function(df, outcome) {
  call <- as.character(match.call())
  ggplot(df, aes(date, {{outcome}})) +
      geom_point() + 
      geom_line() + 
      labs(title = paste("Overall mobility trends in", call[3] ), 
           x = "Date", 
           y = "% change")
}

```

Now let's loop through the data and apply our `create_plot()` function: 

```{r}

plots_by_country_mobil_cat <- by_country_mobil_cat %>%
  mutate(plots = pmap(list(country, mobil_category, data), 
                      ~create_plot(df = by_country_mobil_cat, outcome = .x)))

plots_by_country_mobil_cat$plots[[1]]

# Q for DA: why is this not working? 
```

We can use walk() to save our plots. It's very similar to map() but doesn't print anything to the screen, instead creates files and file pathways to save. 

```{r}

fs::dir_create(here::here("plots", "mobility-plots")) # creates a folder called "plots" and a subdirectory called "mobility"

# then create a vector that has all of the file paths: 

files_c <- str_replace_all(tolower(plots_by_country_mobil_cat$country), " ", "-")
files_m <- str_replace_all(plots_by_country_mobil_cat$mobil_category, "_", "-")
paths <- here::here("plots", "mobility", glue("{files_c}-{files_m}.png"))
paths


walk2(paths, plots_by_country_mobil_cat$plots, ggsave,
      width = 9.5, 
      height = 6.5,
      dpi = 500)

# Q for DA: difference between walk() and walk2()? 

```

## Task 2: How do infection rates relate to mobility trends, by country?

Fitting a linear model to see how mobility changed as infection rate changed in each country: 

```{r results="hide"}

d1 <- d %>%
  mutate(total_mob = rowSums(.[4:9]))

lm_by_country_lists <- d1 %>% 
  split(.$country) %>% # split into separate tbl for each country
  map(~lm(total_mob ~ total_cases, data = .)) # for each, fit lm
# Outputs a nested list

```

Extracting slopes: 

```{r }
slopes <- lm_by_country_lists %>%
    map(coef) %>%
    map_dbl(2)
slopes
```
